Associations between Chlorophyll \emph{a} and various Microcystin-LR Health Advisory Concentrations

### 2015-09-09 manuscript/manuscript.Rmd
new render

---
title: 'Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory Concentrations'
author:  
- name: Jeffrey W. Hollister
  affilnum: 1
  email: hollister.jeff@epa.gov
- name:  Betty J. Kreakie
  affilnum: 1
- name: Dorothy Q. Kellog
  affilnum: 2
affiliation:
- affilnum: 1
  affil: US Environmental Protection Agency, Office of Research and Development, National Health and Environmental Effects Research Laboratory, Atlantic Ecology Division, 27 Tarzwell Drive  Narragansett, RI, 02882, USA
- affilnum: 2
  affil: University of Rhode Island, Department of Natural Resources Science, Kingston, RI, 02882, USA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    template: components/manuscript.latex
  html_document: null
  word_document: null
fontsize: 11pt
capsize: normalsize
csl: components/ecology.csl
documentclass: article
spacing: doublespacing
linenumbers: no
bibliography: components/manuscript.bib
abstract: no
---

@@

In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of Microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie.  Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.  While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.

One of these complications is that they rely on available measurements of microcystin-LR.  This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].  Measurements with greater accuracy and precision requires taking regular water samples and having those samples process in a lab to determine the toxin concentration [REFS].  This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.  Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.

>>>>

---
title: Associations between Chlorophyll *a* and various Microcystin-LR Health Advisory
  Concentrations
author:
- affilnum: 1
  email: hollister.jeff@epa.gov
  name: Jeffrey W. Hollister
- affilnum: 1
  name: Betty J. Kreakie
- affilnum: 2
  name: Dorothy Q. Kellog
affiliation:
- affil: US Environmental Protection Agency, Office of Research and Development, National
    Health and Environmental Effects Research Laboratory, Atlantic Ecology Division,
    27 Tarzwell Drive  Narragansett, RI, 02882, USA
  affilnum: 1
- affil: University of Rhode Island, Department of Natural Resources Science, Kingston,
    RI, 02882, USA
  affilnum: 2
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    template: components/manuscript.latex
  html_document: null
  word_document: null
capsize: normalsize
csl: components/ecology.csl
documentclass: article
fontsize: 11pt
linenumbers: no
bibliography: components/manuscript.bib
spacing: doublespacing
abstract: no
---

@@

In the summer of 2014, the city of Toledo, OH was forced to shut down their municipal water supply due in part to an excess of microcystin-LR that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in Lake Erie [REFS].  Since this event, significant legislation has been passed in the United States and the US Environmental Protection Agency (USEPA) has released suggested microcystin-LR concentrations that would trigger health advisories.  MORE ON THE LEVELS. While these levels and associated advisories are likely to help mitigate the impacts from harmful algal blooms, they are not without complications.

One of these complications is that they rely on available measurements of microcystin-LR.  This toxin can be measured in the field using test strips but these are a coarse measure at best and currently available test strips focus on 1 and 10 $\mu$g/L [REFS].  Measurements with greater accuracy and precision require taking regular water samples and having those samples processed in a lab to determine the toxin concentration [REFS].  This has the potentially to be costly and time consuming both factors which could limit monitoring efforts.  Additionally, microcystin-LR is not currently a routine water quality measurment and thus availability of data will continue to be a problem. Fortunately, microcystin-LR has been shown to be associated with several other, more easily measured components of water quality.

### 2015-09-09 manuscript/manuscript.tex
new render

In the summer of 2014, the city of Toledo, OH was forced to shut down
their municipal water supply due in part to an excess of Microcystin-LR
that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
Lake Erie. Since this event, significant legislation has been passed in
the United States and the US Environmental Protection Agency (USEPA) has
released suggested microcystin-LR concentrations that would trigger
health advisories. While these levels and associated advisories are
likely to help mitigate the impacts from harmful algal blooms, they are
not without complications.

One of these complications is that they rely on available measurements
of microcystin-LR. This toxin can be measured in the field using test
strips but these are a coarse measure at best and currently available
test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
greater accuracy and precision requires taking regular water samples and
having those samples process in a lab to determine the toxin
concentration {[}REFS{]}. This has the potentially to be costly and time
consuming both factors which could limit monitoring efforts.
Fortunately, microcystin-LR has been shown to be associated with several
other, more easily measured components of water quality.

>>>>

In the summer of 2014, the city of Toledo, OH was forced to shut down
their municipal water supply due in part to an excess of microcystin-LR
that resulted from a ongoing cyanobacteria harmful algal bloom (cHAB) in
Lake Erie {[}REFS{]}. Since this event, significant legislation has been
passed in the United States and the US Environmental Protection Agency
(USEPA) has released suggested microcystin-LR concentrations that would
trigger health advisories. MORE ON THE LEVELS. While these levels and
associated advisories are likely to help mitigate the impacts from
harmful algal blooms, they are not without complications.

One of these complications is that they rely on available measurements
of microcystin-LR. This toxin can be measured in the field using test
strips but these are a coarse measure at best and currently available
test strips focus on 1 and 10 \(\mu\)g/L {[}REFS{]}. Measurements with
greater accuracy and precision require taking regular water samples and
having those samples processed in a lab to determine the toxin
concentration {[}REFS{]}. This has the potentially to be costly and time
consuming both factors which could limit monitoring efforts.
Additionally, microcystin-LR is not currently a routine water quality
measurment and thus availability of data will continue to be a problem.
Fortunately, microcystin-LR has been shown to be associated with several
other, more easily measured components of water quality.
